\documentclass[11pt]{article}

% Change "review" to "final" to generate the final (sometimes called camera-ready) version.
% Change to "preprint" to generate a non-anonymous version with page numbers.
\usepackage[preprint]{acl}

% Standard package includes
\usepackage{times}
\usepackage{latexsym}

% For proper rendering and hyphenation of words containing Latin characters (including in bib files)
\usepackage[T1]{fontenc}
% For Vietnamese characters
% \usepackage[T5]{fontenc}
% See https://www.latex-project.org/help/documentation/encguide.pdf for other character sets

% This assumes your files are encoded as UTF8
\usepackage[utf8]{inputenc}

% This is not strictly necessary, and may be commented out,
% but it will improve the layout of the manuscript,
% and will typically save some space.
\usepackage{microtype}

% This is also not strictly necessary, and may be commented out.
% However, it will improve the aesthetics of text in
% the typewriter font.
\usepackage{inconsolata}

% Make links clickable
\usepackage{hyperref}

%Including images in your LaTeX document requires adding
%additional package(s)
\usepackage{graphicx}
\usepackage{listings}

% If the title and author information does not fit in the area allocated, uncomment the following
%
%\setlength\titlebox{<dim>}
%
% and set <dim> to something 5cm or larger.

\title{Smarter Park Search: A RAG Chatbot for Minnesota State Parks}

% Author information can be set in various styles:
% For several authors from the same institution:
% \author{Author 1 \and ... \and Author n \\
%         Address line \\ ... \\ Address line}
% if the names do not fit well on one line use
%         Author 1 \\ {\bf Author 2} \\ ... \\ {\bf Author n} \\
% For authors from different institutions:
% \author{Author 1 \\ Address line \\  ... \\ Address line
%         \And  ... \And
%         Author n \\ Address line \\ ... \\ Address line}
% To start a separate ``row'' of authors use \AND, as in
% \author{Author 1 \\ Address line \\  ... \\ Address line
%         \AND
%         Author 2 \\ Address line \\ ... \\ Address line \And
%         Author 3 \\ Address line \\ ... \\ Address line}

\author{Luca Comba \\
  University of St. Thomas \\ comb6457@stthomas.edu \\}

%\author{
%  \textbf{First Author\textsuperscript{1}},
%  \textbf{Second Author\textsuperscript{1,2}},
%  \textbf{Third T. Author\textsuperscript{1}},
%  \textbf{Fourth Author\textsuperscript{1}},
%\\
%  \textbf{Fifth Author\textsuperscript{1,2}},
%  \textbf{Sixth Author\textsuperscript{1}},
%  \textbf{Seventh Author\textsuperscript{1}},
%  \textbf{Eighth Author \textsuperscript{1,2,3,4}},
%\\
%  \textbf{Ninth Author\textsuperscript{1}},
%  \textbf{Tenth Author\textsuperscript{1}},
%  \textbf{Eleventh E. Author\textsuperscript{1,2,3,4,5}},
%  \textbf{Twelfth Author\textsuperscript{1}},
%\\
%  \textbf{Thirteenth Author\textsuperscript{3}},
%  \textbf{Fourteenth F. Author\textsuperscript{2,4}},
%  \textbf{Fifteenth Author\textsuperscript{1}},
%  \textbf{Sixteenth Author\textsuperscript{1}},
%\\
%  \textbf{Seventeenth S. Author\textsuperscript{4,5}},
%  \textbf{Eighteenth Author\textsuperscript{3,4}},
%  \textbf{Nineteenth N. Author\textsuperscript{2,5}},
%  \textbf{Twentieth Author\textsuperscript{1}}
%\\
%\\
%  \textsuperscript{1}Affiliation 1,
%  \textsuperscript{2}Affiliation 2,
%  \textsuperscript{3}Affiliation 3,
%  \textsuperscript{4}Affiliation 4,
%  \textsuperscript{5}Affiliation 5
%\\
%  \small{
%    \textbf{Correspondence:} \href{mailto:email@domain}{email@domain}
%  }
%}

\begin{document}
\maketitle
\begin{abstract}
The State of Minnesota is home to 65 beautiful state parks. Still, there are few available options for people to search and retrieve information about them. Navigating the available information on the internet can be cumbersome due to the large volume of data and nested web pages. We collected all available data about Minnesota state parks and created a new, optimized way to search for information using Artificial Intelligence and Retrieval-Augmented Generation (RAG) methods. After chunking, indexing, and embedding the data into a vector database, we created a chatbot that efficiently understands users' questions about Minnesota state parks and can accurately respond with valuable information.
\end{abstract}

\section{Introduction}
Minnesota, with its 65 state parks, ranks 15th on the list of United States' state parks by total area, covering 267,000 acres of park land \cite{rankings}. The Minnesota Department of Natural Resources (DNR), through their website \cite{dnr}, publishes general information about each state park, including details on maps, seasonal updates, camping and lodging, events, reservations, trails, recreation facilities, and amenities.

I personally love camping, and since I started a few years ago, I have found it quite difficult to find state park information online. While Google Maps and other third-party tools help inform users through maps and reviews, searching through the vast amount of available data has become difficult and time-consuming.

Retrieval-Augmented Generation (RAG) methods offer an easier way to access and search through a large number of documents. We collected state park data from the internet, cleaned it, and structured it into documents. We then chunked, indexed, and stored the data in a vector database. After preparing the data, we built a chatbot for searching Minnesota State Parks information. This allows users to ask any questions they might have about state parks, and the AI model can formulate an answer based on the collected documents.

\subsection{Minimal Goals}
For our minimal goals, we decided to write a simple scraper to grab content from the web, specifically from the DNR website. We then focused on cleaning, chunking, and indexing the collected data into a vector database. Initially, we thought we could train and fine-tune a Large Language Model (LLM) from scratch, similar to what Andrej Karpathy did with NanoChat \cite{nanochat}. However, due to time and resource constraints, we pivoted to building a RAG solution. This approach retrieves relevant documents in response to a user's question and uses a pre-trained LLM, in our case Gemini \cite{gemini}, to generate an answer and cite the sources. We built a command-line interface (CLI) as the main way to interact with the RAG API. This chat CLI allows the user to type a question, which is then embedded to retrieve related state park documents and passed to the LLM for answer generation.

\subsection{Optimistic Goals}
Our optimistic goals include improving the scraper to collect third-party data, such as photos or Google Maps reviews. We also want to improve the RAG model by adding a re-ranker. Another goal is to build a web interface to allow users to connect and chat directly in their browsers. The deployment of the application would also change, with the application running on a cloud platform with an automated CI/CD pipeline.

\section{Exploring}
Before this class, I had not implemented a full RAG system. I had some experience with web crawling using Selenium \cite{selenium} but not with Puppeteer \cite{puppeteer} and not with end-to-end retrieval pipelines.

Initially, I considered other applications, such as a website that acts as a blueprint and graphical interface for designing AI systems, similar to Figma or LucidChart but with a pre-made list of AI models based on the arXiv dataset and foundational models. I also had an idea for a website that could function as a tool for visually impaired people to help them read emails and navigate the web.

My main sources for topic selection and system design were: class lectures, the course book and the official Minnesota DNR web pages as the project data source.

\section{Building}
The source code for this project is available on GitHub at \href{https://github.com/lukfd/campllm}{https://github.com/lukfd/campllm}.

The goal of the build phase was to create a usable data pipeline for a RAG assistant focused on Minnesota state parks. The core of this pipeline is a set of three command-line interface (CLI) scripts that handle data processing and user interaction.

I used a Puppeteer crawler to collect park content from the Minnesota DNR website. For each park and section, the crawler stores metadata and extracted text. Example metadata fields include park name, park URL, section ID, section heading, and fetch timestamp.

The data processing is orchestrated by three main Python scripts:
\begin{itemize}
    \item \texttt{src/script/index.py}: This script serves as the main entry point for the data pipeline. It can trigger the Node.js crawler to collect fresh data, or it can work with an existing collected data file. The script collects raw data, cleans it, chunks it into smaller sections and uploads them into a ChromaDB \cite{chroma} vector database. The uploaded data is embedded with an embedding model and metadata are stored alongside.
    \item \texttt{src/script/database.py}: This is a utility script for interacting directly with the ChromaDB instance. It lists collections, run semantic queries, or peek at a sample of the stored entries.
    \item \texttt{src/script/chat.py}: This script launches the interactive chatbot. Here, the user asks a question, the script embeds the query, retrieves the most relevant document chunks from ChromaDB, and then passes the question along with the retrieved context to the Gemini LLM \cite{gemini} to generate a answer with source citations.
\end{itemize}

After having collected and processed the data, the cleaned text is chunked into smaller passages. These chunks are uploaded to Chroma and embedded using an embedding model, in our case we used \code{minishlab/potion-base-8M}. The resulting vectors are stored in a vector database: ChromaDB. At query time, the user's question is embedded with the same model. The system then retrieves the top-k most relevant chunks, which can optionally be reranked, before sending the context and the original question to the LLM for answer generation.

\section{Discovering}
Thanks to this project I have learned how to use Puppeteer to crawl and extract data from a website. I have also learned how to use the Chroma database to store and retrieve vector embeddings. Additionally, I have gained experience in building a RAG pipeline and integrating it with a pre-trained LLM for answer generation.

\subsection{Results}
The primary result of this project is a functional RAG chatbot capable of answering questions about Minnesota state parks. The chatbot demonstrates the effectiveness of the data pipeline, from crawling and indexing to retrieval and generation.

The following examples from a chat session illustrate the chatbot's capabilities. In the first interaction, the user asks for specific details about Frontenac State Park:

\begin{lstlisting}[breaklines=true]
You: Hello, tell me about Frontenac state park

CampLLM: Here's what I can tell you about Frontenac 
State Park, based on the provided information:

*   Location & Contact: The park is located at 29223 
    County 28 Boulevard, Frontenac, MN 55026 [Source 3]. 
    You can reach them at 651-299-3000 or 
    frontenac.statepark@state.mn.us [Source 3].
*   Wi-Fi: Wi-Fi is available at the ranger station, 
    and no password is required [Source 3].
*   Firewood: You can purchase firewood at the ranger 
    station when it is open, or pay using Yodel [Source 3]. 
    It is important to only burn approved firewood to 
    protect the trees [Source 3].
*   Campgrounds: Frontenac State Park also manages four 
    campgrounds in the Richard J. Dorer Memorial 
    Hardwood State Forest, which are first-come, 
    first-served [Source 3]. The Kruger Campground has 
    19 campsites [Source 3]. There are also horse 
    camping options at Zumbro Bottoms, generally open 
    May 1 through November 1 [Source 3].

General knowledge: Frontenac State Park is located in 
southeastern Minnesota, overlooking Lake Pepin on the 
Mississippi River. It's known for its bluffs and 
opportunities for hiking and birdwatching.

Sources:
- [Source 3] Frontenac | Directions | 
  https://www.dnr.state.mn.us/state_parks/park.html?id=spk00160#homepage
\end{lstlisting}

This example shows the model successfully retrieving multiple, specific facts from the indexed documents and attributing them to the correct source. It also demonstrates the model's ability to supplement the retrieved information with its own general knowledge.

In the second example, the user asks a broader, location-based question:

\begin{lstlisting}[breaklines=true]
You: What are the state park close to the Twins Cities

CampLLM: Based on the provided information regarding 
Upper Sioux Agency State Park, here's what I can tell 
you about nearby state parks:

*   The information states that Upper Sioux Agency State 
    Park is now closed to the public as of 
    February 16, 2024 [Source 1, 2, 3, 4, 5].
*   There are two state parks within a 45-mile radius 
    of Upper Sioux Agency State Park: Lac Qui Parle and 
    Camden [Source 1, 2, 3, 4, 5].
*   There are three additional state parks within a 
    65-mile radius of Upper Sioux Agency State Park: 
    Monson Lake, Sibley, and [Source 1, 2, 3, 4, 5].

However, this information is *relative to Upper Sioux 
Agency State Park*, and doesn’t directly answer your 
question about parks near the Twin Cities. 

General knowledge: The Twin Cities (Minneapolis and 
St. Paul) are located in southeastern Minnesota. 
State parks relatively close to the Twin Cities include 
Fort Snelling State Park, Interstate State Park, 
William O'Brien State Park, and Afton State Park. 
These are all within roughly an hour's drive, 
depending on traffic.
\end{lstlisting}

This interaction is particularly insightful. The retrieval mechanism found documents related to "parks close to" another park (Upper Sioux Agency), but the LLM correctly identified that this context did not directly answer the user's question about the Twin Cities. It presented the retrieved information transparently, explained its limitations, and then used its general knowledge to provide a more relevant and helpful answer. This highlights the model's ability to reason about the provided context and avoid generating a misleading response.

The chat interface also includes helper commands, such as \texttt{/model} to check the current LLM and \texttt{/q} to exit the session.

\subsection{Self-Assessment}
Before starting this class, I did not have experience with either RAG, vector databases or LLM apis. Thanks to this project I was able to create a working chatbot that can answer questions by utilizing a LLM api and a vector database.

\section{Class Topics}

\subsection{1. Retrieval-Augmented Generation}
I learned that RAG is system for combining document retrieval with generation. Instead of relying on an LLM to answer from its own memory, my project first retrieves relevant text from the park dataset and provides that text to the model as context. This approach helps reduce the risk of hallucination and ensures the answers are grounded in the source documents I collected \cite{rag_lecture}.

\subsection{2. Embedding Models}
An embedding model's job is to convert text into numerical vectors. For this project, it was important that the park document chunks and the user's questions were processed by the same embedding model. This ensures they are comparable in vector space, so that text with similar meanings ends up close together, which is what makes the retrieval process work effectively \cite{embeddings_lecture}.

\subsection{3. Tokenization}
I came to understand that tokenization is a critical step because LLMs and embedding models don't see raw text because they see tokens. \cite{tokenization_lecture}.

\subsection{4. Vector Databases}
A vector database is designed to store these text embeddings and run efficient nearest neighbor searches. For this project, I chose Chroma because it was simple to set up and worked well for local development and testing \cite{vectordb_lecture}.

\subsection{5. Reranking}
The initial retrieval step provides a set of candidate documents, but the top results aren't always the most relevant. Reranking helps improve precision by scoring the candidates a second time using a more sophisticated matching logic. In my project, I implemented a basic reranker and also designed an optional path to use a more advanced cross-encoder reranker \cite{reranking_lecture}.


\section{Limitations}

\subsection{Incomplete Work}
Due to time constraints, I was not able to implement the web interface and automate a build pipeline for the CI/CD pipeline. 

The lack of a more complete dataset that includes more details on the parks, such as photos, reviews, maps, also limits the chatbot's ability to provide complete answers.

The chatbot's prompt is quite basic and could be improved with better instructions, examples, and formatting to help the LLM generate more accurate and informative responses.

\section*{Acknowledgments}
I would like to thank my professor, Dr. Ebrahim “Abe” Kazemzadeh.

I would also like to thank my classmates and the entire University of St. Thomas. 

Thank you to my family, Elena and Albino, and thank you to Gillian, Thomas, Ivan, Tyler and Daniel.

Additially, throughout the project I have utilized Github Copilot and Gemini for code generation and debugging.

% Bibliography entries for the entire Anthology, followed by custom entries
%\bibliography{custom,anthology-overleaf-1,anthology-overleaf-2}

% Custom bibliography entries only
% \bibliography{custom}
\begin{thebibliography}{1}

\bibitem{rankings}
World Population Review.
\newblock State Parks by State.
\newblock \href{https://worldpopulationreview.com/state-rankings/state-parks-by-state}{https://worldpopulationreview.com/state-rankings/state-parks-by-state}.

\bibitem{dnr}
Department of Natural Resources.
\newblock Home Page.
\newblock \href{https://www.dnr.state.mn.us/}{https://www.dnr.state.mn.us/}.

\bibitem{nanochat}
Nanochat.
\newblock Github.
\newblock \href{https://github.com/karpathy/nanochat}{https://github.com/karpathy/nanochat}.

\bibitem{gemini}
Gemini.
\newblock Documentation.
\newblock \href{https://ai.google.dev/gemini-api/docs/models}{https://ai.google.dev/gemini-api/docs/models}.

\bibitem{chroma}
Chroma. A Vector Database.
\newblock \href{https://www.trychroma.com/}{https://www.trychroma.com/}.

\bibitem{selenium}
Selenium.
\newblock \href{https://www.selenium.dev/}{https://www.selenium.dev/}.

\bibitem{puppeteer}
Puppeteer.
\newblock \href{https://pptr.dev/}{https://pptr.dev/}.

\bibitem{campllm}
CampLLM.
\newblock \href{https://github.com/lukfd/campllm}{https://github.com/lukfd/campllm}.

\bibitem{rag_lecture}
Dr. Abe Kazemzadeh.
\newblock Semantic Search and Retrieval-Augmented Generation (RAG).
\newblock \emph{SEIS 767: Conversational AI}, 2026.
\newblock Lecture 8, Page 12.

\bibitem{embeddings_lecture}
Dr. Abe Kazemzadeh.
\newblock Creating Text Embedding Models.
\newblock \emph{SEIS 767: Conversational AI}, 2026.
\newblock Lecture 10, Page 25.

\bibitem{tokenization_lecture}
Dr. Abe Kazemzadeh.
\newblock Tokenization.
\newblock \emph{SEIS 767: Conversational AI}, 2026.
\newblock Lecture 2, Page 16.

\bibitem{vectordb_lecture}
Dr. Abe Kazemzadeh.
\newblock Semantic Search and Retrieval-Augmented Generation (RAG).
\newblock \emph{SEIS 767: Conversational AI}, 2026.
\newblock Lecture 8, Page 18.

\bibitem{reranking_lecture}
Dr. Abe Kazemzadeh.
\newblock Semantic Search and Retrieval-Augmented Generation (RAG).
\newblock \emph{SEIS 767: Conversational AI}, 2026.
\newblock Lecture 8, Page 17.


\end{thebibliography}

\end{document}
